---
title: "p8105_hw1_yw4665"
author: "Yishi Wang"
date: "2025-09-18"
output: github_document
---

### Problem 0.1 & 0.2

All requirements and recommendations from Problem 0.1 and 0.2 are properly followed

### Problem 1

Load the 'moderndive' library:
```{r}
library(moderndive)
```

Load the according dataset. Use 'help()' to see dataset description and variable names:
```{r}
data("early_january_weather")
help("early_january_weather")
```
From the RStudio Help panel, we can see that this dataset includes:

**origin**: Weather station. 

**year, month, day, hour**: Time of recording.

**temp, dewp**: Temperature and dewpoint in F.

**humid**: Relative humidity.

**wind_dir, wind_speed, wind_gust**: Wind direction (in degrees), speed and gust speed (in mph).

**precip**: Precipitation, in inches.

**pressure**: Sea level pressure in millibars.

**visib**: Visibility in miles.

**time_hour**: Date and hour of the recording as a POSIXct date.

Now we use 'nrow()' and 'ncol()' to see the size of the dataset:
```{r}
nrow(early_january_weather)
ncol(early_january_weather)
```

Now we use 'mean()' to see the mean temperature:
```{r}
mean(early_january_weather$temp)
```

Next, we make scatter plot using 'ggplot2' package. First load 'ggplot2':
```{r}
library(ggplot2)
```
Now we create a scatterplot named "plot_time_temp" with proper titles and different colorings based on humid variable. 
```{r}
plot_time_temp = ggplot(
  early_january_weather, 
  aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(title = "Temperature vs Time",
       x = "Time (hour)",
       y = "Temperature (F)",
       )
plot_time_temp
```
From the plot we can see that the temperature forms a clear periodic pattern, where the temperature drops overnight and rises during daytime. Overall, both temperature and humidity increases from new year to January 15th as the spikes are getting larger and the scatters are becoming lighter.

Finally, to save this scatterplot with 'ggsave':
```{r}
ggsave(filename = "temp_vs_time.png", plot = plot_time_temp)
```

### Problem 2


